Millions of Type Errors

### 2023-08-11 programming23/paper.tex
quick related work notes, lots of unbound cites

%% zfstt-ieeesensors-2022

@@

>>>>

%% - zfstt-ieeesensors-2022
%% - lit review, benefits of static types
%%   http://danluu.com/empirical-pl/
%% - thomas kennedy .... generic usage monitoring ascilite 2003
%% - ahmadzadeh elliman higgins iticse 2005
%% - buffardi etal 2014 adaptive and social mechanisms
%% - bddf-icse-2016
%% -  sh edwards and jason snyder icer 2009 comparing effective and ineffective grading platform
%% - murphy etal sigcse 2009 Retina: helping students
%% - ESP workshop on empirical studies of programming
%%   https://dl.acm.org/doi/proceedings/10.1145/266399
%%   nothing sounds big ... empirical yes, large no
%% - Choppella Haynes IU TR 426
%% - Tip Dinesh tosem 2001

\cite{bgimgm-cse-2016}
study Java compiler error messages after an intervention, counting num errors,
counting repeats, counting per user

\cite{anna-russo-kennedy-ms-2006}
BitFit, web-based java ide log num hint requests num, result of check answer
requests num compile, num run, no code?! good we have the same constraint

\cite{ab-sigcse-2015}
frequency, time-to-fix, spread of errors
 (we can't do time)
parsing compile error + source code
 sometimes with custom parsers

\cite{m-masters-2016}
study errors in blackbox dataset,
 overwhelmingly syntax

\cite{bkmu-sigcse-2014}
blackbox, 100k users
source code, diffs
error line, column, message;
keep all comments EXCEPT one above the class declaration

\cite{bask-icer-2018}
2TB data, Java programs
18 publications survey, technical challenges in analysis
 most pubs analyze errors;
nobody asked:
 which exns do students encounter (what?! mccall kolling 2014)
 execution / editing timelines
(blackbox analysis tutorial in sigcse 2020 on a mini dataset, to help researchers start writing an analysis)

\cite{t-hatra-2021}
plan for a large user-centered study 
... not very deep, look at edit sequences, replay for errors

\cite{cdhhjklwya-hatra-2020}
evaluating the PLIERS design process (Coblenz etal 202X)
teaching students to use PLIERS; everyone ran a user study!
6 languages
 3 ran usability studies
recommend talk-alouds

\cite{gstf-hatra-2021}
liquid types for java
30 devs user study
 zoom

\cite{lfgc-pldi-2007}
typechecker never errors,
 calls a helper that looks for nearby programs that do typecheck
tested on
 10 students, 2yrs professional experience
 5 hw assignments
 2122 files collected;
won \pct{19}, lost \pct{17};
ill-suited for things like (let x = e1 in ....),
 change e1,
 use typechecker feedback to guide edits;
small data!

\cite{w-popl-1986}
reporting source, not detection
tested on 9 type errors
 all \emph{deletions}, such as forgetting to inject in an enum

\cite{sscwj-oopsla-2017}
NATE numerical analysis of type errors
5000 labeled programs
 errors from class, first fix

\cite{sjw-jfp-2018}
type error witness generation
4500 ill-typed student programs
 \pct{85} of time, synthesis works
 \pct{70} of time, locate source

\cite{h-dissertiation-2005}
constraint-based type inference;
general te criteria:
- target audience (novice / expert)
- output format (text / viz)
- interactivity (yes / no)
- heuristics (yes / no)
- primary-or-external (most type errors are easy, some could really use dedicated help)

\cite{hw-scp-2004}
location of type error as a slice,
 not a srcloc,
 not a subtree

@@

TODO Liblit dissertiation
